---
title: "SVD_Final"
author: "Ihor Markevych"
date: "12/7/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read data

```{r}
d <- as.matrix(read.csv('imageSVD.csv', header=F))
```

# Functions

## Cumulative variance

```{r}
cumVar <- function(A){
   eigs <- svd(A)$d ^2
   return(cumsum(eigs / sum(eigs)))
}
```

## Compression

Bellow I use built in `svd()` function.

```{r}
compress <- function(d, k){
    svd <- svd(d, nu=k, nv=k)
    sigma <- svd$d
    U <- svd$u
    V <- svd$v
    Ak <- matrix(rep(0, len=dim(d)[1]^2), nrow=dim(d)[1])
    for(i in 1:k){
        Ak <- Ak + sigma[i] * U[,i] %*% t(V[,i])
    }
    return(Ak)
}
```

# Question(a)

```{r}
minNumToPreserve <- which(cumVar(d)>=0.9)[1]
```

So, `r minNumToPreserve` singular values/vectors are needed to preserve 90% of variance of the original image.

# Question (b)

```{r}
plot(cumVar(d), xlab='k', ylab='Cummulative varience')
```

# Question (c)

```{r}
Ak <- compress(d, minNumToPreserve)
```

**Pixels:**   

$A_k (1,1)=$ `r Ak[1,1]`.  
$A_k (256,256)=$ `r Ak[256,256]`.  
$A_k (512,512)=$ `r Ak[512,512]`.

# Question (d)

```{r}
err <- norm(d - Ak, type="2") 
varLost <- 1 - cumVar(d)[minNumToPreserve]
```

Error from built-in function is `r err`. This is $L_2$-norm (Euclidian), which is $\sqrt{\sum A_{ij} - (A_k)_{ij}}$.  
Error from singular values can be calculated as variance lost, which is `r varLost * 100`%.


# Question (e)

## Original
```{r}
mean <- 123.55
rotate <- function(x) t(apply(x, 2, rev))

image(rotate(d) + mean, useRaster=TRUE, axes=FALSE, col = grey(seq(0, 1, length = 256)))
```

## Compressed

```{r}
image(rotate(Ak)+mean, useRaster=TRUE, axes=FALSE, col = grey(seq(0, 1, length = 256)))
```