---
title: "SVD_NumberRecognition"
author: "Ihor Markevych"
date: "12/4/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mcompanion)
library(pracma)
```

# Functions

## SVD

### Singular values

Computation of singular values of a matrix.

```{r}
singularValues <- function(A){
    eigs <- eigen(t(A) %*% A)$values 
    return(sqrt(eigs[eigs > 0]))
}
```

### SVD itself

Full SVD decomposition.

```{r}
SVD <- function(A){
    sV <- singularValues(A)
    sigma <- diag(sV, nrow = dim(A)[1], ncol = dim(A)[2])
    
    V <- eigen(t(A) %*% A)$vectors
    
    U <- A %*% V[, 1] / sV[1]
    for(i in 2:length(sV)){
        U <- cbind(U, A %*% V[, i] / sV[i])
    }  
    U <- cbind(U, null_complement(U))
    
    return(list('U' = U, 'Sigma' = sigma, 'V' = V))
    
}
```

## Vector operations

### Dot product

Dot product in $R^n$.

```{r}
dot <- function(x, y){
  sum(x * y)
}
```

### L2 Norm 

$L^2$ norm in $R^n$.

```{r}
norm2 <- function(x) {
  sqrt(dot(x, x))
}
```

## Least Squares

### Solution

$\hat{x}$ in Least Squares Model.

```{r}
xHat <- function(A, b) {
  inv(t(A) %*% A) %*% t(A) %*% b
}
```

### Error of a model

Error of Least Squares Model.

```{r}
error <- function(A,x,b) {
  norm2(A %*% x - b)
}
```

### Create a model and calculate error

Next function creates a least squares model and calculates its error from matrix `X` and vector `y`.

```{r}
leastSquaresError <- function(X, y) {
  y <- as.matrix(y)
  X <- as.matrix(X)
  sol <- xHat(X, y)
  return(error(X, sol, y))
}
```

## Image displaying

Display image in grayscale.

```{r}
displayImage <- function(im){
    rotate <- function(x) t(apply(x, 2, rev))
    im <- matrix(im, sqrt(length(im)), sqrt(length(im)), byrow=T)
    image(rotate(im), useRaster=TRUE, axes=FALSE, col = grey(seq(1, 0, length = 256)))
}
```


# Data reading

Train and test data reading.

```{r}
trainInput <- t(read.csv('trainInput.csv', header=F))
trainOutput <- c(read.csv('trainOutput.csv', header=F))
testInput <- t(read.csv('testInput.csv', header=F))
testOutput <- c(read.csv('testOutput.csv', header=F))
```

Create a `digits` variable that will containall digits that are present in a dataset (for scenario where we may not have one digits, e.g. `7`).

```{r}
digits <- unique(trainOutput)
```

Separate train dataset by output labels.

```{r}
digitTrain <- list()
for(i in 1:length(digits)){
    digitTrain[[i]] <- na.omit(as.matrix(trainInput[trainOutput == digits[[i]],]))
}
names(digitTrain) <- digits
```

# Visual investigation

## Train data

Display first 8 images from train data.

```{r}
par(mfrow=c(2,4))
for(i in 1:8){
    displayImage(trainInput[i,])
}
```

## Display first 8 images from test data

Display first 8 images from test data.

```{r}
par(mfrow=c(2,4))
for(i in 1:8){
    displayImage(testInput[i,])
}
```

## Singular images

### Function

Calculare and display singular image number `imageNumber` of digit `digit` (both string and integer are taken by function).

```{r}
displaySingularImage <- function(digit, imageNumber=1){
    im <- SVD(digitTrain[[toString(digit)]])$V
    im <- im[, imageNumber]
    displayImage(im)
}
```

### Singular image 1

Display first singular images of each digit.

```{r}
par(mfrow=c(3,4))
for(i in 0:9){
    displaySingularImage(i)
}
```

### Singular image 5

Display singular image number 5 of each digit.

```{r}
par(mfrow=c(3,4))
for(i in 0:9){
    displaySingularImage(i, 5)
}
```

### Singular image 100

Display singular image number 100 of each digit. Those images contains a mostly noise.

```{r}
par(mfrow=c(3,4))
for(i in 0:9){
    displaySingularImage(i, 100)
}
```

# Training model

## Model

Create a a model by taking first 20 singular images to represent each digit.

```{r}
model <- list()

for(i in 1:length(digits)){
    model[[i]] <- SVD(digitTrain[[toString(digits[i])]])$V[,1:20]
}
names(model) <- digits
```

## Predict function

Predict digit of input vector by finding least squares model (based on 20 singular images of each digit) with smallest residual.

```{r}
predict <- function(y){
    errors <- c()
    for(i in digits){
        errors <- c(errors, leastSquaresError(model[[toString(i)]], y))
    }
    return(digits[which(errors == min(errors))])
}
```

# Testing model

Create a vector of predicted labels.

```{r}
predictions <- c()
for(i in 1:nrow(testInput)){
    predictions <- c(predictions, predict(testInput[i,]))
}
```

Create vector `names` that contains sorted digits that are present in a dataset.

```{r}
names <- unlist(digits)
names <- sort(names)
names <- lapply(names, toString)
```

## Confusion matrix

Function that calculates elements of confusion matrix.

```{r}
confusion <- function(i,j){
    return(sum(predictions[testOutput==i] == j))
}
```

Create confusion matrix.

```{r}
confusionMatrix <- matrix(nrow = length(digits), ncol = length(digits))

rownames(confusionMatrix) <- names
colnames(confusionMatrix) <- names

for(i in names){
    for(j in names){
        confusionMatrix[[i,j]] <- confusion(i,j)
    }   
}
```

In this confusion matrix:  
rownames are actual labels; 
colnames - predicted labels.  
  
`ftable` is used todisplay dimension names.   

```{r}
dimnames(confusionMatrix) <- list(actual=rownames(confusionMatrix), predicted=colnames(confusionMatrix))

ftable(confusionMatrix, row.vars='actual', col.vars='predicted')
```

## Accuracy for each digit

### Recall

**Recall** refers to the percentage of total relevant results correctly classified by algorithm.

$$precision=\frac{TruePositive}{TruePositive+FalseNegative}$$

```{r}
recall <- c()
for(i in names){
    recall <- c(recall, confusionMatrix[i,i] / sum(confusionMatrix[,i]))
}

for(i in 1:length(names)){
    cat(cat('For digit', names[[i]], 'recall score is', recall[i]), '.\n', sep='')
}
```

### Precision

*Precision** means the percentage of your results which are relevant 

$$precision=\frac{TruePositive}{TruePositive+FalsePositive}$$

```{r}
precision <- c()
for(i in names){
    precision <- c(precision, confusionMatrix[i,i] / sum(confusionMatrix[i,]))
}

for(i in 1:length(names)){
    cat(cat('For digit', names[[i]], 'precision score is', precision[i]), '.\n', sep='')
}
```

### F1

F1-score is the harmonic mean of precision and recall.  

$$F1=2*\frac{Precision*Recall}{Precision+Recall}$$

```{r}
f1 <- 2 * precision * recall / (precision + recall)

for(i in 1:length(names)){
    cat(cat('For digit', names[[i]], 'F1 score is', f1[i]), '.\n', sep='')
}
```

## Overall	correct	classification	rate

Rate how many samples we classified correctly to total number of samples.

```{r}
sum(diag(confusionMatrix)) / sum(confusionMatrix) * 100
```